%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
                                                          % if you need a4paper
%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
                                                          % paper

\IEEEoverridecommandlockouts                              % This command is only
                                                          % needed if you want to
                                                          % use the \thanks command
\overrideIEEEmargins
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{import}
\usepackage{lettrine}
\usepackage{nameref}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{pdfpages}
\usepackage[margin=0.5in]{geometry} % for PAPER & MARGIN
\usepackage[many]{tcolorbox} % for COLORED BOXES (tikz and xcolor included)
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\newtheorem{definition}{Definition}[section]
\graphicspath{{./images/}}
\definecolor{main}{HTML}{DADADA}    % setting main color to be used
\definecolor{sub}{HTML}{F5F5F5}

\newtcolorbox{boxK}{
    sharpish corners, % better drop shadow
    boxrule = 0pt,
    toprule = 4.5pt, % top rule weight
    enhanced,
    fuzzy shadow = {0pt}{-2pt}{-0.5pt}{0.5pt}{black!35} % {xshift}{yshift}{offset}{step}{options} 
}

\newtcolorbox{boxF}{
    colback = sub,
    enhanced,
    boxrule = 1.5pt, 
    colframe = white, % making the base for dash line
    borderline = {0.8pt}{0pt}{main, dashed} % add "dashed" for dashed line
}


% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
Beyond Tokens: A Zero-Trust AuthZ Protocol for the Agentic Era
}

%\author{ \parbox{3 in}{\centering Huibert Kwakernaak*
%         \thanks{*Use the $\backslash$thanks command to put information here}\\
%         Faculty of Electrical Engineering, Mathematics and Computer Science\\
%         University of Twente\\
%         7500 AE Enschede, The Netherlands\\
%         {\tt\small h.kwakernaak@autsubmit.com}}
%         \hspace*{ 0.5 in}
%         \parbox{3 in}{ \centering Pradeep Misra**
%         \thanks{**The footnote marks may be inserted manually}\\
%        Department of Electrical Engineering \\
%         Wright State University\\
%         Dayton, OH 45435, USA\\
%         {\tt\small pmisra@cs.wright.edu}}
%}

\author{Nicola Gallo$^{1}$% <-this % stops a space
\thanks{$^{1}$Nicola Gallo, Software Architect at Nitro Agility. 
This work was carried out as part of professional activities within Nitro Agility and should be understood in that professional context. 
{\tt\small nicola.gallo@nitroagility.com}}%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
\import{./sections}{abstract.tex}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction}
\import{./sections}{introduction.tex}

\section{Tokens in the Authorization Perspective}
\label{sec:token-authz-prospective}
\import{./sections}{token-authz-prospective.tex}

\section{The Industry Evolution to Zero Trust}
\label{sec:the-industry-evolution-zerotrust}
\import{./sections}{the-industry-evolution-zerotrust}


\section*{Acknowledgments}
This text was originally written by the author in English. 
As a non-native English speaker, AI-based writing tools were employed solely to refine grammar, clarity, and fluency in order to speed up the writing process and reduce language errors.
All ideas, arguments, and original contributions are entirely the author's own.

\vspace{1em}
The views and proposals expressed in this paper are intended for discussion and research purposes only. 
They should not be interpreted as finalized standards or production-ready specifications.

\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{thebibliography}{99}
\bibitem{c1} Brewer, Eric. (2000). Towards robust distributed systems. PODC. 7. 10.1145/343477.343502. 
\bibitem{c2} V. Velepucha and P. Flores, "A Survey on Microservices Architecture: Principles, Patterns and Migration Challenges," in IEEE Access, vol. 11, pp. 88339-88358, 2023, doi: 10.1109/ACCESS.2023.3305687.
\bibitem{c3} V. Vernon, Domain-Driven Design Distilled. Reading, MA, USA: Addison-Wesley Professional, 2016.
\bibitem{c4} [OpenID.2.0] OpenID Foundation, “OpenID Authentication 2.0,” December 2007.
\bibitem{c5} [RFC6749]	M. Jones, D. Hardt, “The OAuth 2.0 Authorization Framework,” October 2012.
\bibitem{c6} [RFC6750]	M. Jones, D. Hardt, “The OAuth 2.0 Authorization Framework: Bearer Token Usage", October 2012
\bibitem{c7} L. Zhi, W. Jing, C. Xiao-su and J. Lian-xing, "Research on Policy-based Access Control Model," 2009 International Conference on Networks Security, Wireless Communications and Trusted Computing, Wuhan, China, 2009, pp. 164-167, doi: 10.1109/NSWCTC.2009.313.
\bibitem{c8} Nicola Gallo, Antonio Radesca, "A Multi-Account and Multi-Tenant Policy-Based Access Control (PBAC) Approach for Distributed Systems Augmented with Risk Scores Generation".
\bibitem{c9} D. J. Abadi, “Consistency Tradeoffs in Modern Distributed Database System Design: CAP is Only Part of the Story,” \emph{IEEE Computer}, vol. 45, no. 2, pp. 37–42, Feb. 2012. DOI: 10.1109/MC.2012.33
\end{thebibliography}

\end{document}