\lettrine{T}{he} \textbf{Agentic Era} is emerging in industry discussions, as many recognize that certain assumptions no longer hold and past trade-offs are becoming increasingly fragile.  
Change can often be unsettling, especially when existing systems and business models are built on established approaches. 
Yet it may be time to start with a blank page and re-examine the compromises we have inherited.

\vspace{0.5em} The community should take this as a challenge to rethink the \textbf{foundations of authorization}, especially in the context of distributed systems and \textbf{AI agents}.

\vspace{0.5em} In the near future, companies will continue to rely on \textbf{OAuth} to authenticate humans, and on solutions such as \textbf{WIMSE} or \textbf{SPIFFE} to authenticate non-human workloads.  
Yet at some point, we must stop patching legacy models and acknowledge that \textbf{new architectures demand new foundations}.

\vspace{0.5em} This new foundation can open entirely new markets. 
It can still leverage existing technologies, but with a fresh approach that enables \textbf{innovation}, \textbf{new value creation}, and the development of novel products.  
Industry has been built around \textbf{identities}, and most importantly this new foundation could enable the real adoption of \textbf{AI agents} within the enterprise.

\vspace{0.5em} The \textbf{authentication layer} is relatively solid, whether using centralized or decentralized approaches, but \textbf{authorization} still works only under certain \textbf{trade-offs}.  
These trade-offs cannot support AI agents and, more generally, distributed systems. 

\vspace{0.5em} \textbf{History of science} teaches us that progress often comes from looking across disciplines, borrowing solutions, and applying them in new contexts, that is what this paper aims to do looking at \textbf{Enterprise Patterns} for distributed systems.  
It is also important to consider this in the context of \textbf{Zero Trust Network Access (ZTNA)} and, more broadly, the \textbf{Zero Trust model}, which will be crucial for the future of secure access.

\vspace{0.5em} This paper adopts such an approach, challenging the core trade-off that the industry has relied on: the \textbf{identity- and HTTP-centric model}.  
An \textbf{AI agent protocol} would need to include many other features; however, this paper focuses specifically on \textbf{authorization (authz)}.
